---
title: "Batch Effect Correction_Chip"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 120),
  message = FALSE,
  warning = FALSE
)
```
# load libraries
```{r}
library(devtools)
library(dplyr)
library(tidyr)
library(Seurat)
library(ggplot2)
library(readxl)
library(harmony)
library(scran)
library(SeuratWrappers)
library(SeuratData)
library(liger)
library(kBET)
library(FNN)
library(lisi)
library(magrittr)
library(limma)
library(WeightedCluster)
library(cluster)
library(mclust)
library(batchelor)
library(caret)
library(ggpubr)
library(AnnotationHub)
library(SingleR)
library(celldex)
```

# Load Data
```{r init}
# load library info
lib.info <- read_xlsx('H:\\GSAT Rotation1\\data\\library_info_5_and_3.xlsx')
# load data for individual samples and merge together
data <- NULL
for (i in 1:nrow(lib.info)) {
i.df <- lib.info[i,]
i.dir <- paste0(i.df$AnalysisDir)
i.data <- Read10X(data.dir = i.dir)
i.obj <- CreateSeuratObject(counts = i.data, project = i.df$StudyID)
i.obj <- RenameCells(i.obj, add.cell.id = i.df$StudyID)
if (is.null(data)) { data <- i.obj } else { data <- merge(data, i.obj)}
}
# Add annotation about chip (batch)
data@meta.data <- cbind(
data@meta.data, 
lib.info[match(data$orig.ident, lib.info$StudyID), 
c("StudyID", "Chip", "Type", 'Flowcell','Assay')])

# check data
data
```
Result:
merged data:
36601 features across 31358 samples within 1 assay 

# Standard pre-processing workflow
```{r mito, fig.height=7, fig.width=13}
# add mt gene percent as a column
data[["percent.mt"]] <- PercentageFeatureSet(object = data, pattern = "^MT-")
```
## Filter Cells
keep cells which meet these requirements
1. mt gene < 10% 
2. nfeature <3000
3. nfeature >300  
3. ncounts < 13000
```{r qc2, fig.height=7, fig.width=13}
# visualize QC metrics as a violin plot
VlnPlot(object = data, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3, group.by = 'StudyID')

# visualize the distribution of counts detected per cell via histogram
data@meta.data %>% 
  	ggplot(aes(color=orig.ident, x=nCount_RNA, fill= orig.ident)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	scale_x_log10() + 
  	geom_vline(xintercept = 300)

# visualize the distribution of genes detected per cell via histogram
data@meta.data %>% 
  	ggplot(aes(color=orig.ident, x=nFeature_RNA, fill= orig.ident)) + 
  	geom_density(alpha = 0.2) + 
  	theme_classic() +
  	scale_x_log10() + 
  	geom_vline(xintercept = 300)

# visualize feature-feature relationships
plot1 <- FeatureScatter(object = data, feature1 = "nCount_RNA", feature2 = "percent.mt", group.by = 'StudyID') 
plot2 <- FeatureScatter(object = data, feature1 = "nCount_RNA", feature2 = "nFeature_RNA", group.by = 'StudyID') 
CombinePlots(plots = list(plot1,plot2))

# visualize the relationship between nFeature and nCount via scatter plot
data@meta.data %>% 
  	ggplot(aes(x=nCount_RNA, y=nFeature_RNA, color=percent.mt)) + 
  	geom_point() + 
	  scale_colour_gradient(low = "gray90", high = "black") +
  	stat_smooth(method=lm) +
  	scale_x_log10() + 
  	scale_y_log10() + 
  	theme_classic() +
  	geom_vline(xintercept = 500) +
  	geom_hline(yintercept = 250) +
  	facet_wrap(~orig.ident)

# subset data 
data <- subset(x = data, subset = nFeature_RNA > 300 & nFeature_RNA < 3000 & percent.mt < 10 & nCount_RNA <15000)
data
```
result:
In the raw dataset, we have 31358 cells.
After filtering, we have 27719 cells (88% of the raw cells are kept).

## Normalization
```{r normalize}
# data normalization
data <- NormalizeData(object = data, normalization.method = "LogNormalize", scale.factor = 1e4)
```
## HVG
```{r var_features, fig.height=5, fig.width=11}
# find highly variable genes( n = 2000)
data <- FindVariableFeatures(object = data,selection.method = 'vst', nfeatures = 2000)

# check top 10 highly variable genes
top10 <- head(x = VariableFeatures(object = data), 10)

top10
# plot variable features
plot1 <- VariableFeaturePlot(object = data)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
CombinePlots(plots = list(plot1, plot2))
plot2
```

# PCA
## RunPCA
```{r pca,results='hide'}
# scaling data
data <- ScaleData(data, vars.to.regress = "percent.mt")
# run pca
data <- RunPCA(object = data, features = VariableFeatures(object = data))
```
## PCA Visualization
```{r multi-heatmap, fig.height=12, fig.width=9}
DimPlot(object = data, reduction = 'pca', group.by = 'Chip')
DimHeatmap(object = data, dims = 1:15, cells = 500, balanced = TRUE)
```

# Batch Effect Correction
## Harmony
Procedure: Merged 3' and 5' dataset, QC, normalization, find HVGs, scaling, PCA, RunHarmony(), clustering, runUMAP()
The Vignette can be found at: https://htmlpreview.github.io/?https://github.com/satijalab/seurat.wrappers/blob/master/docs/harmony.html
```{r tsneplot, fig.height=5, fig.width=13}
# run harmony 
data.harmony <- RunHarmony(data, group.by.vars = 'Chip', reduction = 'pca')

# clustering
data.harmony<-  data.harmony %>%   
    FindNeighbors(reduction = "harmony", dims = 1:40) %>% 
    FindClusters(resolution = 1.5) %>% 
    RunUMAP(reduction = "harmony", dims = 1:40) 

# visualize ump result
harmony_plot <-DimPlot(data.harmony, reduction = "umap", label = TRUE, pt.size = .1, group.by = c('ident','Chip', 'StudyID'))
harmony_plot
```
## FastMNN
Procedure: Merged 3' and 5' dataset, QC, normalization, find HVGs, RunFastMNN(), clustering, runUMAP()
The Vignette can be found at: https://htmlpreview.github.io/?https://github.com/satijalab/seurat.wrappers/blob/master/docs/fast_mnn.html
```{r tsneplot, fig.height=5, fig.width=13}
# create a new seurat object that is not being processed
data1 <- NULL
for (i in 1:nrow(lib.info)) {
i.df <- lib.info[i,]
i.dir <- paste0(i.df$AnalysisDir)
i.data <- Read10X(data.dir = i.dir)
i.obj <- CreateSeuratObject(counts = i.data, project = i.df$StudyID)
i.obj <- RenameCells(i.obj, add.cell.id = i.df$StudyID)
if (is.null(data1)) { data1 <- i.obj } else { data1 <- merge(data1, i.obj)}
}
# Add annotation about chip (batch)
data1@meta.data <- cbind(
data1@meta.data, 
lib.info[match(data1$orig.ident, lib.info$StudyID), 
c("StudyID", "Chip", "Type", 'Flowcell','Assay')])
# add mt percent
data1[["percent.mt"]] <- PercentageFeatureSet(object = data1, pattern = "^MT-")
# subset data
data1 <- subset(x = data1, subset = nFeature_RNA > 300 & nFeature_RNA < 3000 & percent.mt < 10 & nCount_RNA <15000)

# normalization and find HVGs
data.mnn<- NormalizeData(data1, normalization.method = "LogNormalize", scale.factor = 1e4) %>%
           FindVariableFeatures(selection.method = 'vst', nfeatures = 5000)

# run Fast MNN based on Chip type
data.mnn <- RunFastMNN(object.list = SplitObject(data.mnn, split.by = 'Chip'))

# run UMAP
data.mnn <- data.mnn %>%
   FindNeighbors(reduction = 'mnn', dims = 1:40) %>%
   FindClusters(resolution = 1.5) %>%
   RunUMAP(reduction = "mnn", dims = 1:40) 

# visualize by individual samples
mnn_plot <- DimPlot(data.mnn, reduction = "umap", label = TRUE, pt.size = .1, group.by = c('ident','Chip', 'StudyID'))
mnn_plot
```
## Seurat
Procedure: Individual dataset, QC, normalization, integration, scaling, pca,  clustering, runUMAP()
The Vignette can be found at: https://satijalab.org/seurat/v3.2/integration.html
```{r tsneplot, fig.height=5, fig.width=13}
# separate merged data into individual datasets based on Chip type
data.list <- SplitObject(data1, split.by = 'Chip')
# perform data preprocessing for individual samples
for (i in 1:length(x = data.list)) {
    data.list[[i]] <- NormalizeData(object = data.list[[i]], verbose = FALSE)
    data.list[[i]] <- FindVariableFeatures(object = data.list[[i]], 
        selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}
# find anchors for performing CCA
data.anchors <- FindIntegrationAnchors(object.list = data.list, dims = 1:40)
# integrate datasets 
data.seurat <- IntegrateData(anchorset = data.anchors, dims = 1:40) %>%
               ScaleData(verbose = FALSE) %>%
               RunPCA(npcs = 40, verbose = FALSE) %>%
               FindNeighbors(reduction = "pca", dims = 1:40) %>%
               FindClusters(resolution = 1.5) %>%
               RunUMAP(reduction = "pca", dims = 1:40)
# visualize the results
seurat_plot <- DimPlot(data.seurat, reduction = "umap", label = TRUE, pt.size = .1, group.by = c('ident','Chip', 'StudyID'), ncol = 3)
seurat_plot
```
##LIGER
Procedure: Merged 3' and 5' dataset, QC, normalization, find HVGs, scaling, RunOptimizeALS(), RunQuantileNorm(), clustering, runUMAP()
The Vignette can be found at: https://htmlpreview.github.io/?https://github.com/satijalab/seurat.wrappers/blob/master/docs/liger.html
```{r tsneplot, fig.height=5, fig.width=13}
# normalization, find HVGs, and scaling
## data1 is already subset but not normalized
data.liger <- NormalizeData(data1, normalization.method = "LogNormalize", scale.factor = 1e4) %>%
              FindVariableFeatures(selection.method = 'vst', nfeatures = 2000) %>%
              ScaleData(split.by = 'Chip', do.center = FALSE)

data.liger <- RunOptimizeALS(data.liger, k = 20, lambda = 5, split.by = 'Chip') %>%
              RunQuantileNorm(split.by = "Chip")

data.liger <- FindNeighbors(data.liger, reduction = "iNMF", dims = 1:20) %>%
              FindClusters(resolution = 1.5) %>%
              RunUMAP(dims = 1:ncol(data.liger[["iNMF"]]), reduction = "iNMF")

# visualize ump result
liger_plot <- DimPlot(data.liger, reduction = "umap", label = TRUE, pt.size = .1, group.by = c('ident','Chip', 'StudyID'), ncol = 3)
liger_plot
```
## Raw
### Clustering
```{r}
data <- FindNeighbors(data, dims = 1: 40) %>%
        FindClusters(resolution = 1.5) %>%
        RunUMAP(dims = 1:40)
```
### Visualization
```{r tsneplot, fig.height=5, fig.width=15}
raw_plot <- DimPlot(data, reduction = 'umap', group.by = c('ident','Chip','StudyID'), label = T)
raw_plot
```

# Evaluate Batch Effect
## Extract corrected matrixs from Seurat Object
```{r}
matrix.raw <- data@reductions$pca@cell.embeddings
matrix.harmony <- data.harmony@reductions$harmony@cell.embeddings
matrix.mnn <- data.mnn@reductions$mnn@cell.embeddings
matrix.seurat <- data.seurat@reductions$pca@cell.embeddings
matrix.liger <-data.liger@reductions$iNMF@cell.embeddings
```

## kBET
### Raw
```{r}
batch <- as.factor(data$Chip)
# calculate k0 and knn
k0 <- floor(mean(table(batch)))
knn.raw <- get.knn(matrix.raw, k = k0, algorithm = 'cover_tree')
# calculate kBet based on different testsize
batch.estimate.raw_5 <- kBET(matrix.raw, batch = batch, k0=k0, knn = knn.raw, verbose = TRUE, testSize = 5)
batch.estimate.raw_10 <- kBET(matrix.raw, batch = batch, k0=k0, knn = knn.raw, verbose = TRUE, testSize = 10)
batch.estimate.raw_15 <- kBET(matrix.raw, batch = batch, k0=k0, knn = knn.raw, verbose = TRUE, testSize = 15)
batch.estimate.raw_20 <- kBET(matrix.raw, batch = batch, k0=k0, knn = knn.raw, verbose = TRUE, testSize = 20)
batch.estimate.raw_25 <- kBET(matrix.raw, batch = batch, k0=k0, knn = knn.raw, verbose = TRUE, testSize = 25)
```
### Harmony
```{r}
# calculate knn
knn.harmony <- get.knn(matrix.harmony, k = k0, algorithm = 'cover_tree')
# calculate kBet based on different testsize
batch.estimate.harmony_5 <- kBET(matrix.harmony, batch = batch, k0=k0, knn = knn.harmony, verbose = TRUE, testSize = 5)
batch.estimate.harmony_10 <- kBET(matrix.harmony, batch = batch, k0=k0, knn = knn.harmony, verbose = TRUE, testSize = 10)
batch.estimate.harmony_15 <- kBET(matrix.harmony, batch = batch, k0=k0, knn = knn.harmony, verbose = TRUE,  testSize = 15)
batch.estimate.harmony_20 <- kBET(matrix.harmony, batch = batch, k0=k0, knn = knn.harmony, verbose = TRUE,  testSize = 20)
batch.estimate.harmony_25 <- kBET(matrix.harmony, batch = batch, k0=k0, knn = knn.harmony, verbose = TRUE,  testSize = 25)
```
### Fast MNN
```{r}
# calculate knn
knn.mnn <- get.knn(matrix.harmony, k = k0, algorithm = 'cover_tree')
# calculate kBet based on different testsize
batch.estimate.mnn_5 <- kBET(matrix.mnn, batch = batch,k0=k0, knn = knn.mnn,verbose = TRUE, testSize = 5)
batch.estimate.mnn_10 <- kBET(matrix.mnn, batch = batch, k0=k0, knn = knn.mnn, verbose = TRUE, do.pca = FALSE, testSize = 10)
batch.estimate.mnn_15 <- kBET(matrix.mnn, batch = batch, k0=k0, knn = knn.mnn, verbose = TRUE,  testSize = 15)
batch.estimate.mnn_20 <- kBET(matrix.mnn, batch = batch, k0=k0, knn = knn.mnn, verbose = TRUE,  testSize = 20)
batch.estimate.mnn_25 <- kBET(matrix.mnn, batch = batch, k0=k0, knn = knn.mnn, verbose = TRUE,  testSize = 25)
```
### Seurat
```{r}
# calculate knn
knn.seurat <- get.knn(matrix.seurat, k = k0, algorithm = 'cover_tree')
# calculate kBet based on different testsize
batch.estimate.seurat_5 <- kBET(matrix.seurat, batch = batch, k0=k0, knn = knn.seurat, verbose = TRUE, testSize = 5)
batch.estimate.seurat_10 <- kBET(matrix.seurat, batch = batch, k0=k0, knn = knn.seurat,verbose = TRUE)
batch.estimate.seurat_15 <- kBET(matrix.seurat, batch = batch, k0=k0, knn = knn.seurat,verbose = TRUE,  testSize = 15)
batch.estimate.seurat_20 <- kBET(matrix.seurat, batch = batch, k0=k0, knn = knn.seurat,verbose = TRUE,  testSize = 20)
batch.estimate.seurat_25 <- kBET(matrix.seurat, batch = batch, k0=k0, knn = knn.seurat,verbose = TRUE,  testSize = 25)
```
### LIGER
```{r}
# calculate knn
knn.liger <- get.knn(matrix.liger, k = k0, algorithm = 'cover_tree')
# calculate kBet based on different testsize
batch.estimate.liger_5 <- kBET(matrix.liger, batch = batch, k0=k0, knn = knn.liger, verbose = TRUE, testSize = 5)
batch.estimate.liger_10 <- kBET(matrix.liger, batch = batch, k0=k0, knn = knn.liger,verbose = TRUE)
batch.estimate.liger_15 <- kBET(matrix.liger, batch = batch, k0=k0, knn = knn.liger,verbose = TRUE,  testSize = 15)
batch.estimate.liger_20 <- kBET(matrix.liger, batch = batch, k0=k0, knn = knn.liger,verbose = TRUE,  testSize = 20)
batch.estimate.liger_25 <- kBET(matrix.liger, batch = batch, k0=k0, knn = knn.liger,verbose = TRUE,  testSize = 25)
```

## LISI
### Raw
```{r}
# calculate iLISI score for the uncorrected dataset
ilisi_raw <- compute_lisi(matrix.raw, meta_data = data@meta.data, label_colnames = 'Chip')
# calculate median value of iLISI score
ilisi_raw_median <- median(ilisi_raw$Chip)

# calculate cLISI score for the uncorrected dataset
clisi_raw <- compute_lisi(matrix.raw, meta_data = data@meta.data, label_colnames = 'seurat_clusters') 
# calculate median value of clisi score            
clisi_raw_median <- median(clisi_raw$seurat_clusters)
```

### Harmony
```{r}
# calculate iLISI score
ilisi_harmony <- compute_lisi(matrix.harmony, meta_data = data.harmony@meta.data, label_colnames = 'Chip')
# calculate median value of iLISI score
ilisi_harmony_median <- median(ilisi_harmony$Chip)

# calculate cLISI score and normalize the data
clisi_harmony <- compute_lisi(matrix.harmony, meta_data = data.harmony@meta.data, label_colnames = 'seurat_clusters') # calculate median value of clisi score  
clisi_harmony_median <- median(clisi_harmony$seurat_clusters)
```
### Fast MNN
```{r}
# calculate iLISI score 
ilisi_mnn <- compute_lisi(matrix.mnn, meta_data = data.mnn@meta.data, label_colnames = 'Chip')
# calculate median value of iLISI score
ilisi_mnn_median <- median(ilisi_mnn$Chip)

# calculate cLISI score and normalize the data
clisi_mnn <- compute_lisi(matrix.mnn, meta_data = data.mnn@meta.data, label_colnames = 'seurat_clusters') 
# calculate median value of clisi score 
clisi_mnn_median <- median(clisi_mnn$seurat_clusters)
```
### Seurat
```{r}
# calculate iLISI score
ilisi_seurat <- compute_lisi(matrix.seurat, meta_data = data.seurat@meta.data, label_colnames = 'Chip')
# calculate median value of iLISI score
ilisi_seurat_median <- median(ilisi_seurat$Chip)

# calculate cLISI score 
clisi_seurat <- compute_lisi(matrix.seurat, meta_data = data.seurat@meta.data, label_colnames = 'seurat_clusters') 
# calculate median value of clisi score
clisi_seurat_median <- median(clisi_seurat$seurat_clusters)
```
### LIGER
```{r}
# calculate iLISI score
ilisi_liger <- compute_lisi(matrix.liger, meta_data = data.liger@meta.data, label_colnames = 'Chip')
# calculate median value of iLISI score
ilisi_liger_median <- median(ilisi_liger$Chip)

# calculate cLISI score and normalize the data
clisi_liger <- compute_lisi(matrix.liger, meta_data = data.liger@meta.data, label_colnames = 'seurat_clusters') 
# calculate cLISI score 
clisi_liger_median <- median(clisi_liger$seurat_clusters)
```

## ASW
### Raw
```{r}
# calculate euclidean distance in the matrix
matrix.raw_dist <- dist(matrix.raw)

# calculate asw score for batch effects 
asw_raw_batch <- wcSilhouetteObs(matrix.raw_dist, clustering = data$Chip, measure = 'ASW')
asw_raw_batch_median <- median(asw_raw_batch)

# calculate asw score for cell purity
asw_raw_cell <- wcSilhouetteObs(matrix.raw_dist, clustering = data$seurat_clusters, measure = 'ASW')  
asw_raw_cell_median <- median(asw_raw_cell)
```
### Harmony
```{r}
# calculate euclidean distance in the matrix
matrix.harmony_dist <- dist(matrix.harmony)

# calculate asw score for batch effects 
asw_harmony_batch <- wcSilhouetteObs(matrix.harmony_dist, clustering = data.harmony$Chip, measure = 'ASW')
asw_harmony_batch_median <- median(asw_harmony_batch)

# calculate asw score for cell purity
asw_harmony_cell <- wcSilhouetteObs(matrix.harmony_dist, clustering = data.harmony$seurat_clusters, measure = 'ASW')
asw_harmony_cell_median <- median(asw_harmony_cell)
```
### Fast MNN
```{r}
# calculate euclidean distance in the matrix
matrix.mnn_dist <- dist(matrix.mnn)

# calculate asw score for batch effects 
asw_mnn_batch <- wcSilhouetteObs(matrix.mnn_dist, clustering = data.mnn$Chip, measure = 'ASW')
asw_mnn_batch_median <- median(asw_mnn_batch)

# calculate asw score for cell purity
asw_mnn_cell <- wcSilhouetteObs(matrix.mnn_dist, clustering = data.mnn$seurat_clusters, measure = 'ASW') 
asw_mnn_cell_median <- median(asw_mnn_cell)
```
### Seurat
```{r}
# calculate euclidean distance in the matrix
matrix.seurat_dist <- dist(matrix.seurat)

# calculate asw score for batch effects
asw_seurat_batch <- wcSilhouetteObs(matrix.seurat_dist, clustering = data.seurat$Chip, measure = 'ASW')
asw_seurat_batch_median <- median(asw_seurat_batch)

# calculate asw score for cell purity
asw_seurat_cell <- wcSilhouetteObs(matrix.seurat_dist, clustering = data.seurat$seurat_clusters, measure = 'ASW') 
asw_seurat_cell_median <- median(asw_seurat_cell)
```
### LIGER
```{r}
# calculate euclidean distance in the matrix
matrix.liger_dist <- dist(matrix.liger)

# calculate asw score for batch effects
asw_liger_batch <- wcSilhouetteObs(matrix.liger_dist, clustering = data.liger$Chip, measure = 'ASW')
asw_liger_batch_median <- median(asw_liger_batch)

# calculate asw score for cell purity
asw_liger_cell <- wcSilhouetteObs(matrix.liger_dist, clustering = data.liger$seurat_clusters, measure = 'ASW') 
asw_liger_cell_median <- median(asw_liger_cell)
```

## ARI
### Raw
```{r}
# calculate kmeans
k.raw_cell <- kmeans(matrix.raw, centers = length(levels(data$seurat_clusters)))
# calculate ari score for cell type purity
ari_raw_cell <- adjustedRandIndex(as.numeric(data$seurat_clusters), as.numeric(k.raw_cell$cluster))

# calculate kmeans
k.raw_batch <- kmeans(matrix.raw, centers = 3)
cluster_batch <- data@meta.data %>%
     mutate(cluster = case_when(Chip == "CHIP013" ~ 3,
                                Chip == "CHIPCCC" ~ 1,
                                Chip == "CHIPDDD" ~ 2))
# calculate ari score for batch
ari_raw_batch <- adjustedRandIndex(as.numeric(cluster_batch$cluster), as.numeric(k.raw_cell$cluster))

```
### Harmony
```{r}
# calculate kmeans
k.harmony_cell <- kmeans(matrix.harmony, centers = length(levels(data.harmony$seurat_clusters)))
# calculate ari score for cell type purity
ari_harmony_cell <- adjustedRandIndex(as.numeric(data.harmony$seurat_clusters), as.numeric(k.harmony_cell$cluster))

# calculate kmeans
k.harmony_batch <- kmeans(matrix.harmony, centers = 3)
# calculate ari score for batch
ari_harmony_batch <- adjustedRandIndex(as.numeric(cluster_batch$cluster), as.numeric(k.harmony_cell$cluster))

```
### Fast MNN
```{r}
# calculate kmeans
k.mnn_cell <- kmeans(matrix.mnn, centers = length(levels(data.mnn$seurat_clusters)))
# calculate ari score for cell type purity
ari_mnn_cell <- adjustedRandIndex(as.numeric(data.mnn$seurat_clusters), as.numeric(k.mnn_cell$cluster))

# calculate kmeans
k.mnn_batch <- kmeans(matrix.mnn, centers = 3)
# calculate ari score for cell type purity
ari_mnn_batch <- adjustedRandIndex(as.numeric(cluster_batch$cluster), as.numeric(k.mnn_cell$cluster))
```
### Seurat
```{r}
# calculate kmeans
k.seurat_cell <- kmeans(matrix.seurat, centers = length(levels(data.seurat$seurat_clusters)))
# calculate ari score for cell type purity
ari_seurat_cell <- adjustedRandIndex(as.numeric(data.seurat$seurat_clusters), as.numeric(k.seurat_cell$cluster))

# calculate kmeans
k.seurat_batch <- kmeans(matrix.seurat, centers = 3)
# calculate ari score for cell type purity
ari_seurat_batch <- adjustedRandIndex(as.numeric(cluster_batch$cluster), as.numeric(k.seurat_cell$cluster))
```
### LIGER
```{r}
# calculate kmeans
k.liger_cell <- kmeans(matrix.liger, centers = length(levels(data.liger$seurat_clusters)))
# calculate ari score for cell type purity
ari_liger_cell <- adjustedRandIndex(as.numeric(data.liger$seurat_clusters), as.numeric(k.liger_cell$cluster))

# calculate kmeans
k.liger_batch <- kmeans(matrix.liger, centers = 3)
# calculate ari score for cell type purity
ari_liger_batch <- adjustedRandIndex(as.numeric(cluster_batch$cluster), as.numeric(k.liger_cell$cluster))
```

# Save result into a datatable
## LISI
```{r}
# store LISI results into a datatable
lisi_mnn_result <- c(ilisi_mnn_median, clisi_mnn_median) 
lisi_harmony_result <- c(ilisi_harmony_median, clisi_harmony_median)
lisi_seurat_result <- c(ilisi_seurat_median, clisi_seurat_median)
lisi_liger_result <- c(ilisi_liger_median, clisi_liger_median)
lisi_raw_result <- c(ilisi_raw_median, clisi_raw_median) 
lisi_result <- rbind(lisi_raw_result, lisi_harmony_result, lisi_mnn_result, lisi_seurat_result, lisi_liger_result)
colnames(lisi_result) <- c('iLISI_median','cLISI_median')
lisi_result <- lisi_result %>%
               as_tibble() %>%
               mutate(reverse_cLISI = 1 - cLISI_median,
                      iLISI_norm = (iLISI_median - min(iLISI_median)) / (max(iLISI_median) - min(iLISI_median)),
                      cLISI_norm = (cLISI_median - min(cLISI_median)) / (max(cLISI_median) - min(cLISI_median)),
                      fscore = 2 * (1 - cLISI_norm) * iLISI_norm / (1 - cLISI_norm + iLISI_norm))
rownames(lisi_result) <- c('Raw','Harmony','FastMNN','Seurat','LIGER')
```
## ASW
```{r}
# store ASW results into a datatable
asw_mnn_result <- c(asw_mnn_batch_median, asw_mnn_cell_median) 
asw_harmony_result <- c(asw_harmony_batch_median, asw_harmony_cell_median)
asw_seurat_result <- c(asw_seurat_batch_median, asw_seurat_cell_median)
asw_liger_result <- c(asw_liger_batch_median, asw_liger_cell_median)
asw_raw_result <- c(asw_raw_batch_median, asw_raw_cell_median) 
asw_result <- rbind(asw_raw_result, asw_harmony_result, asw_mnn_result, asw_seurat_result, asw_liger_result)
colnames(asw_result) <- c('ASW_batch_median','ASW_cell_median')

asw_result <- asw_result %>%
              as_tibble() %>%
              mutate(ASW_batch_norm = (ASW_batch_median - min(ASW_batch_median)) / (max(ASW_batch_median) - min(ASW_batch_median)),
                     ASW_cell_norm = (ASW_cell_median - min(ASW_cell_median)) / (max(ASW_cell_median) - min(ASW_cell_median)),
                     fscore = (2 * (1 - ASW_batch_norm) *ASW_cell_norm) / (1- ASW_batch_norm + ASW_cell_norm),
                     reverse_batch_median = 1 - ASW_batch_median)
rownames(asw_result) <- c('Raw','Harmony','FastMNN','Seurat','LIGER')

```
## ARI
```{r}
# store ARI result into a datatable
ari_mnn_result <- c(ari_mnn_batch, ari_mnn_cell) 
ari_harmony_result <- c(ari_harmony_batch, ari_harmony_cell)
ari_seurat_result <- c(ari_seurat_batch, ari_seurat_cell)
ari_liger_result <- c(ari_liger_batch, ari_liger_cell)
ari_raw_result <- c(ari_raw_batch, ari_raw_cell) 
ari_result <- rbind(ari_raw_result, ari_harmony_result, ari_mnn_result, ari_seurat_result, ari_liger_result)
colnames(ari_result) <- c('ARI_batch_median','ARI_cell_median')
ari_result <- ari_result %>%
              as_tibble() %>%
              mutate(ARI_batch_norm = (ARI_batch_median - min(ARI_batch_median)) / (max(ARI_batch_median) - min(ARI_batch_median)),
                     ARI_cell_norm = (ARI_cell_median - min(ARI_cell_median)) / (max(ARI_cell_median) - min(ARI_cell_median)),
                     fscore = (2 * (1 - ARI_batch_norm) *ARI_cell_norm) / (1- ARI_batch_norm + ARI_cell_norm),
                     reverse_ari_batch = 1 - ARI_batch_median)
rownames(ari_result) <- c('Raw','Harmony','FastMNN','Seurat','LIGER')
```

## Visualize result
```{r}
LISI_plot <-lisi_result %>%
            ggplot(aes(x = reverse_cLISI, y = iLISI_median, label = rownames(lisi_result))) +
            geom_point(aes(colour = rownames(lisi_result))) +
            geom_text(aes(colour = rownames(lisi_result)), hjust = 0, nudge_x = -0.01) +
            xlab('1 - cLISI cell type') +
            ylab('iLISI batch') +
            labs(color = 'Methods')+
            theme(axis.line = element_line(colour = "black"),
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(),
                  panel.border = element_blank(),
                  panel.background = element_blank()) 
LISI_plot

ASW_plot <- asw_result %>%
            ggplot(aes(x = ASW_cell_median, y = reverse_batch_median, label = rownames(asw_result))) +
            geom_point(aes(colour = rownames(asw_result))) +
            geom_text(aes(colour = rownames(asw_result)), hjust = 0, nudge_x = -0.0045) +
            xlab('ASW cell type') +
            ylab('1 -ASW batch') +
            labs(color = 'Methods') +
            theme(axis.line = element_line(colour = "black"),
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(),
                  panel.border = element_blank(),
                  panel.background = element_blank()) 
ASW_plot

ARI_plot <- ari_result %>%
            ggplot(aes(x = ARI_cell_median, y = reverse_ari_batch, label = rownames(ari_result))) +
            geom_point(aes(colour = rownames(ari_result))) +
            geom_text(aes(colour = rownames(ari_result)), hjust = 0, nudge_x = -0.015) +
            xlab('ARI cell type') +
            ylab('1 -ARI batch') +
            labs(color = 'Methods') +
            theme(axis.line = element_line(colour = "black"),
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(),
                  panel.border = element_blank(),
                  panel.background = element_blank()) 
ARI_plot

evaluation_plot <- ggarrange(LISI_plot, ASW_plot, ARI_plot, common.legend = T, legend = 'right')
evaluation_plot
```

# Assign Cell Types
```{r}
# get reference data
ref_data <- HumanPrimaryCellAtlasData()
# pull matrix from Seurat Object
singler_data <- as.SingleCellExperiment(data)
# Get cell type information
cell <- SingleR(singler_data, ref_data, labels = ref_data$label.main)
# Store cell type information into Seurat Object
data[['SingleR.cluster.labels']] <- cell$labels
data.harmony[['SingleR.cluster.labels']] <- cell$labels
data.mnn[['SingleR.cluster.labels']] <- cell$labels
data.seurat[['SingleR.cluster.labels']] <- cell$labels
data.liger[['SingleR.cluster.labels']] <- cell$labels
```
